---
title: "Model Checks (Cont.)"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}

# Working script, subject to change

library(pROC) # For ROC/AUC

options(scipen = 999)

#source(here::here("setup_environment.R")) # For libraries and filepaths
source("/conf/linkage/output/charlh07/hsmr-pra/setup_environment.R")

detach("package:tidylog", unload=TRUE)

data <- read.csv(paste0(data_folder, # base files with predprob
                        pub_day,
                        "/base_files/", 
                        pub_day, 
                        "_SMR-with-predprob.csv"))

loc <- read.csv(paste0(here::here(), "/reference_files/location_lookups.csv"))

```

NB: Data is from the `r pub_day` publication.

## Decision Thresholds

### Overview

The first element of the HSMR modelling process is a logistic regression model.

A regression model is a model that tries to fit a mathematical function to a set of data. The output of the model is a number;
this number is usually in the range 0-1 and can be treated as a probability. For the HSMR model, this is the probability that a patient will die within 30 days of admission, based on their characteristics.

Once we know these probabilities, we can use them to calculate the number of patients we expect to die within 30 days of admission.
We do this by summing up all of the output probabilities from the model.

For example, imagine a group of 100 patients, all with a probability of death of 0.2 - we would expect (100x0.2 = 20) deaths. If 50 of
these patients had a probability of death of 0.1 instead, we would expect (50x0.1 + 50x0.2 = 15) deaths. If all 100 had a slightly
different probability of death, we would expect (p1 + p2 + .... + p99 + p100) deaths.

There is another type of machine learning model - a 'classification' model. A classification model tries to decide which group
a particular piece of data falls into, usually out of two different groups.

For the HSMR model, this means turning the output - a probability between 0 and 1 - into a YES or NO answer to the question,
"will this patient die within 30 days of admission?"

Turning a regression model into a classification model is as simple as deciding on a 'decision threshold' - a value between 0-1,
where any number below the threshold can be taken as "NO" and any number equal to, or above, the threshold is "YES".

The simplest example of a decision threshold is 0.5 - if a number is less than 0.5 we would round it down to 0, and at 0.5 or over
we would round up to 1. In this case, 0 = NO and 1 = YES, so a higher probability will translate into a YES.

### Method

In a simple classification problem, we would be able to choose the decision threshold ourselves. For the HSMR model this is not the case.

This is because we already know how many YES outcomes we should see, because we have already calculated the number of expected deaths
by summing up the output probabilities from the model. Instead, we need to find the decision threshold that leads to this number
of expected deaths, or YES outcomes.

We need to find the threshold T such that the number of output probabilities greater than, or equal to, T, is the same as the sum of 
all the output probabilities.

We can't solve this analytically (ie. using a function to get the exact answer). Instead, we need to use a minimisation algorithm.

This means that we trial different values of T until we find one that satisfies the condition.

### Derivation

For each geography level (HB, hospital, Scotland) we have the following known values:

-   Number of observed deaths, $N_o$

-   Model output pred_eq $p_i$ for each patient $i$

-   Number of expected deaths, $N_e = \Sigma p_i$

Using these, we aim to define the threshold probability $p_x$ for which:

$$ f(p_i) = \begin{cases} \mbox{true} & \mbox{if } p_i \geq p_x \\ \mbox{false} & \mbox{otherwise} \end{cases} $$ where $f(p_i)$ is a prediction of death within 30 days of admission. This can also be stated with binary output rather than boolean, as:

$$ f(p_i) = \begin{cases} \mbox{1} & \mbox{if } p_i \geq p_x \\ \mbox{0} & \mbox{otherwise} \end{cases} $$ Recall that we know the value of $N_e$, which indicates the number of `true` outcomes we expect to see. This means that:

$$ \sum_i f(p_i)  \approx N_e \therefore \sum_i f(p_i)  \approx \sum_i p_i $$

With the key inequality being that $f(p_i)$ is integer-valued, and $p_i$ is not. As such, $f(p_i) \neq p_i$ and $\sum_i f(p_i) \neq N_e$.

To rephrase this programmatically, we want to find `x` such that `sum(p >= x) - round(sum(p)) == 0` for a vector `p` of values `p[i]`.

From the inequality above, we know that this cannot be solved analytically. This means it must be solved iteratively, using a minimisation algorithm to trial values of `x` until we find one to satisfy the stated condition.

Because the function $f(p_i)$ does not have a continuous derivative, we will need to define our own minimisation algorithm instead of using an out-of-the-box optimiser from R.

## Code

```{r func_defs}

# Define p, a vector of model output pred_eq values.

p <- data$pred_eq


# Define the function f, as described above.

f <- function(p, x) {
  
  # sum_p   = sum of pred_eq values
  # count_p = number of pred_eq values above threshold x
  # out     = function output; sum_p - count_p
  
  sum_p <- round(sum(p))
  
  count_p <- sum(p >= x)
  
  out <- sum_p - count_p
  
  return(out)
  
}


# Define the minimiser function minf, to minimise f with respect to x.

# There are three arguments to this function:
# p   = a vector of pred_eq values p[i] 
# x0  = an initial guess for the value of x
# inc = the base increment to change x by

# Smaller values of inc are more likely to give an accurate answer, but will
# make the loop take longer to run.

minf <- function(p, x0 = 0.5, inc = 1e-5){
  
  out <- 1              # dummy value for output of function f
  it <- 0               # number of iterations
  x <- x0               # current value of x
  
  
  while(out != 0){      # loop breaks when output of function f == 0
    
    out <- f(p, x)      # get output of function f
    
    # we want to change the value of x until we find one that gives the result
    # f(p, x) == 0. We could do this just by adding or subtracting our increment
    # value inc, but inc is set to be very small, so this could take a long
    # time!
    
    # to speed things up, we will define a multiplier or scale factor, a.
    # The value of a should be larger when we are further away from 0, and
    # smaller when we get closer. This lets us make bigger jumps in the value
    # of x when we know that we are not near the answer.
    
    a <- 10^(floor(log10(sqrt(out**2)))-2)
    
    # this means that if the current value of out is an order of magnitude N,
    # we should scale up our increment by an order of N-2.
    
    # If the value of out is +ve, then we need to make x smaller. If out
    # is -ve, we need to make x bigger.
    # Either way, we do this by an increment of (inc * a).
    
    ifelse(out < 0,
          x <- x + (inc * a),
          x <- x - (inc * a))
    
    it <- it + 1 # increase the number of iterations by 1
    
    # every 100th iteration:
    
    if(it %% 100 == 0){
      if(x == x0){ # if x unchanged in 100 iterations
        print(paste0("Best guess: x = ", x, " to accuracy ", inc))
        break # break the loop
      }
    x0 <- x # record current value of x
    }
  
  } # close while() loop

  if(out == 0){
    print(paste0("Solved! x = ", x))
  }

  return(x)
  
}
```

We run the function `minf(p, x)` using the `pred_eq` values for the whole of Scotland, and a first estimate of `x0 = 0.5`, for the default increment size of 0.0001.

```{r run_funcs}
x <- minf(p, 0.5)
```

To verify this result, we look at the `pred_eq` values in the dataset.

The sum of all `pred_eq` values is `r sum(p)`.

The number of `pred_eq` values greater than, or equal to, `x`, is `r sum(p > x)`.

For the whole of Scotland, for the full dataset, the probability threshold $p_x$ is `r round(x, 8)`.

## Application

In reality, we calculate the SMR for the whole of Scotland and for each hospital separately.

> The threshold $p_x$ is different for every location.

We can calculate these thresholds as demonstrated above.

```{r loop_funcs, eval = FALSE}

# EDIT: read back through the code, and 'period' means YEAR not quarter.

# Tidy up location codes, and summarise

grp <- data %>%
  mutate(location = case_when(
      location == "C206H" ~ "C418H",
      location == "G207H" ~ "G107H",
      location == "G306H" ~ "G405H",
      location == "G516H" ~ "G405H",
      location == "Y104H" ~ "Y146H",
      location == "R101H" ~ "R103H",
      TRUE ~ location)) %>%
  group_by(location, period) %>%
  summarise() %>%
  mutate(threshold = 0) # 124 rows total

# Loop over every group

for(i in 1:nrow(grp)){
  group <- grp[i,]
  p <- left_join(group, data)$pred_eq  # get vector of pred_eq for this group
  x <- minf(p, 0.5)                    # evaluate threshold
  grp[i, 3] <- x                       # store result
  print(i)
} 
```

To avoid rerunning the above loop, the threshold information has been saved out previously.

```{r save_temp, include=FALSE, eval=FALSE}
temp_out <- data %>%
  mutate(location = case_when(
      location == "C206H" ~ "C418H",
      location == "G207H" ~ "G107H",
      location == "G306H" ~ "G405H",
      location == "G516H" ~ "G405H",
      location == "Y104H" ~ "Y146H",
      location == "R101H" ~ "R103H",
      TRUE ~ location)) %>%
  full_join(grp)

saveRDS(temp_out,
        paste0("/conf/quality_indicators/hsmr/projects/roc/",
               pub_day,
               "_SMR-with-thresholds-annual.rds"))
```

We can verify the results as we did in the above section.

```{r sanity_check, message=FALSE}

results <- readRDS(paste0("/conf/quality_indicators/hsmr/projects/roc/",
                   pub_day,
                   "_SMR-with-thresholds-annual.rds"))

results_check <- results %>%
                 group_by(location, period, threshold) %>%
                 summarise(preds_above_threshold = sum(pred_eq >= threshold),
                           sum_preds             = round(sum(pred_eq)))

fail <- results_check %>% filter(preds_above_threshold != sum_preds)

print(fail)

```

Of the `r nrow(results_check)` hospitals, there are `r nrow(fail)` for which the threshold could not be calculated exactly.

This could be because the increment accuracy was too low - we can rerun these cases with a smaller increment to see if that helps.

```{r small_inc, message=FALSE}

for(i in 1:nrow(fail)){
  group <- fail[i,]
  p <- left_join(group, data)$pred_eq
  x <- minf(p, group$threshold, inc=1e-8)
  fail[i, 3] <- x
}
 
```

Using an increment of $1e^{-8}$, all of these were solvable.

> We now have threshold $p_x$ values for all hospitals.

```{r save_temp_again, include=FALSE, eval=FALSE}

fail_check <- fail %>%
  left_join(data)  %>%
  group_by(location, quarter, new_threshold = threshold) %>%
  summarise(preds_above_threshold = sum(pred_eq >= new_threshold),
            sum_preds             = round(sum(pred_eq)))

results <- readRDS(paste0("/conf/quality_indicators/hsmr/projects/roc/",
                   pub_day,
                   "_SMR-with-thresholds-annual.rds"))

tst <- fail_check %>%
  select(location, quarter, new_threshold) %>%
  right_join(results) %>%
  mutate(threshold = case_when(!is.na(new_threshold) ~ new_threshold,
                               .default = threshold)) %>%
  select(-new_threshold)

saveRDS(tst,
        paste0("/conf/quality_indicators/hsmr/projects/roc/",
               pub_day,
               "_SMR-with-thresholds-annual_final.rds"))
```

```{r read_final, include=FALSE, message=FALSE}

results <- readRDS(paste0("/conf/quality_indicators/hsmr/projects/roc/",
                   pub_day,
                   "_SMR-with-thresholds-annual_final.rds"))
```

## Results

### True and False Positives

There are 4 possible categories for a predicted outcome to fall into; true positive, false positive, true negative and false negative. These are summarised in the table below.

|                |                    |                     |
|----------------|--------------------|---------------------|
| True Positive  | Prediction = TRUE  | Observation = TRUE  |
| False Positive | Prediction = TRUE  | Observation = FALSE |
| True Negative  | Prediction = FALSE | Observation = FALSE |
| False Negative | Prediction = FALSE | Observation = TRUE  |

In this case, TRUE represents a death within 30 days of admission.

```{r check_results, include=FALSE}

results_check <- results %>%
  mutate(obs = case_match(death30,
                          1 ~ TRUE,
                          .default = FALSE),
         pred = case_when(pred_eq < threshold ~ FALSE,
                          TRUE ~ TRUE)) %>% 
  group_by(location, obs, pred) %>%
  summarise(N = n())

results_check_scot <- results_check %>%
  group_by(obs, pred) %>%
  summarise(N = sum(N)) %>%
  mutate(location = "Scot")

results_check <- rbind(results_check, results_check_scot) %>%
  left_join(loc)

named_locations <- results_check %>% 
  filter(!is.na(location_name)) %>%
  group_by(location) %>%
  mutate(pcent = round(100*N/sum(N), 2))

saveRDS(named_locations,
        paste0("/conf/quality_indicators/hsmr/projects/roc/",
               pub_day,
               "_confusion-data.rds"))
```