---
title: "Model Checks (Cont.)"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}

library(pROC) # For ROC/AUC

options(scipen = 999)

#source(here::here("setup_environment.R")) # For libraries and filepaths
source("/conf/linkage/output/charlh07/hsmr-pra/setup_environment.R")

data <- read.csv(paste0(data_folder, # base files with predprob
                        pub_day,
                        "/base_files/", 
                        pub_day, 
                        "_SMR-with-predprob.csv"))

```

### Decision Thresholds - Overview

For each geography level (HB, hospital, Scotland) we have the following known values:

-   Number of observed deaths, $N_o$

-   Model output pred_eq $p_i$ for each patient $i$

-   Number of expected deaths, $N_e = \Sigma p_i$

Using these, we aim to define the threshold probability $p_x$ for which:

$$ f(p_i) = \begin{cases} \mbox{true} & \mbox{if } p_i \geq p_x \\ \mbox{false} & \mbox{otherwise} \end{cases} $$
where $f(p_i)$ is a prediction of death within 30 days of admission. This can also be stated with binary output rather than boolean, as:

$$ f(p_i) = \begin{cases} \mbox{1} & \mbox{if } p_i \geq p_x \\ \mbox{0} & \mbox{otherwise} \end{cases} $$
Recall that we know the value of $N_e$, which indicates the number of `true` outcomes we expect to see. This means that:

$$ \sum_i f(p_i)  \approx N_e \therefore \sum_i f(p_i)  \approx \sum_i p_i $$

With the key inequality being that $f(p_i)$ is integer-valued, and $p_i$ is not. As such, $f(p_i) \neq p_i$ and $\sum_i f(p_i) \neq N_e$.

To rephrase this programmatically, we want to find `x` such that `sum(p >= x) - round(sum(p)) == 0` for a vector `p` of values `p[i]`.

From the inequality above, we know that this cannot be solved analytically. This means it must be solved iteratively, using a minimisation algorithm to trial values of `x` until we find one to satisfy the stated condition.

Because the function $f(p_i)$ does not have a continuous derivative, we will need to define our own minimisation algorithm instead of using an out-of-the-box optimiser from R.

### Methodology - Overview

```{r func_defs}

# Define p, a vector of model output pred_eq values.

p <- data$pred_eq


# Define the function f, as described above.

f <- function(p, x) {
  
  # sum_p   = sum of pred_eq values
  # count_p = number of pred_eq values above threshold x
  # out     = function output; sum_p - count_p
  
  sum_p <- round(sum(p))
  
  count_p <- sum(p >= x)
  
  out <- sum_p - count_p
  
  return(out)
  
}


# Define the minimiser function minf, to minimise f with respect to x.

# There are three arguments to this function:
# p   = a vector of pred_eq values p[i] 
# x0  = an initial guess for the value of x
# inc = the base increment to change x by

# Smaller values of inc are more likely to give an accurate answer, but will
# make the loop take longer to run.

minf <- function(p, x0, inc = 1e-5){
  
  out <- 1              # dummy value for output of function f
  it <- 0               # number of iterations
  x <- x0               # current value of x
  
  
  while(out != 0){      # loop breaks when output of function f == 0
    
    out <- f(p, x)      # get output of function f
    
    # we want to change the value of x until we find one that gives the result
    # f(p, x) == 0. We could do this just by adding or subtracting our increment
    # value inc, but inc is set to be very small, so this could take a long
    # time!
    
    # to speed things up, we will define a multiplier or scale factor, a.
    # The value of a should be larger when we are further away from 0, and
    # smaller when we get closer. This lets us make bigger jumps in the value
    # of x when we know that we are not near the answer.
    
    a <- 10^(floor(log10(sqrt(out**2)))-2)
    
    # this means that if the current value of out is an order of magnitude N,
    # we should scale up our increment by an order of N-2.
    
    # If the value of out is +ve, then we need to make x smaller. If out
    # is -ve, we need to make x bigger.
    # Either way, we do this by an increment of (inc * a).
    
    ifelse(out < 0,
          x <- x + (inc * a),
          x <- x - (inc * a))
    
    it <- it + 1 # increase the number of iterations by 1
    
    # every 100th iteration:
    
    if(it %% 100 == 0){
      if(x == x0){ # if x unchanged in 100 iterations
        print(paste0("Best guess: x = ", x, " to accuracy ", inc))
        break # break the loop
      }
    x0 <- x # record current value of x
    }
  
  } # close while() loop

  if(out == 0){
    print(paste0("Solved! x = ", x))
  }

  return(x)
  
}
```

We run the function `minf(p, x)` using the `pred_eq` values for the whole of Scotland, and a first estimate of `x0 = 0.5`, for the default increment size of 0.0001.

```{r run_funcs}
x <- minf(p, 0.5)
```


To verify this result, we look at the `pred_eq` values in the dataset.

The sum of all `pred_eq` values is `r sum(p)`.

The number of `pred_eq` values greater than, or equal to, `x`, is `r sum(p > x)`.

For the whole of Scotland, for the full dataset, the probability threshold $p_x$ is `r round(x, 8)`.

### Application of Methodology



#### Granularity

In reality, we calculate the SMR for the whole of Scotland and for each hospital separately, and we do it for each period included in the publication. This means that each of these cohorts will have a separate threshold $p_x$.


