---
title: "Model Checks (Decision Threshold)"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}

library(pROC) # For ROC/AUC

#source(here::here("setup_environment.R")) # For libraries and filepaths
source("/conf/linkage/output/charlh07/hsmr-pra/setup_environment.R")

data <- read.csv(paste0(data_folder, # base files with predprob
                        pub_day,
                        "/base_files/", 
                        pub_day, 
                        "_SMR-with-predprob.csv"))

```

### Method

For each geography level (HB, hospital, Scotland) we have the following known values:

-   Number of observed deaths, $N_o$

-   Model output pred_eq $p_i$ for each patient $i$

-   Number of expected deaths, $N_e = \Sigma p_i$

Using these, we aim to define the threshold probability $p_x$ for which:

$$ f(p_i) = \begin{cases} \mbox{true} & \mbox{if } p_i \geq p_x \\ \mbox{false} & \mbox{otherwise} \end{cases} $$
where $f(p_i)$ is a prediction of death within 30 days of admission. This can also be stated with binary output rather than boolean, as:

$$ f(p_i) = \begin{cases} \mbox{1} & \mbox{if } p_i \geq p_x \\ \mbox{0} & \mbox{otherwise} \end{cases} $$
Recall that we know the value of $N_e$, which indicates the number of `true` outcomes we expect to see. This means that:

$$ \sum_i f(p_i)  \approx N_e \therefore \sum_i f(p_i)  \approx \sum_i p_i $$

With the key inequality being that $f(p_i)$ is integer-valued, and $p_i$ is not. As such, $f(p_i) \neq p_i$ and $\sum_i f(p_i) \neq N_e$.

To rephrase this programmatically, we want to find `x` such that `sum(p >= x) - round(sum(p)) == 0` for a vector `p` of values `p[i]`.

From the inequality above, we know that this cannot be solved analytically. This means it must be solved iteratively, using a minimisation algorithm to trial values of `x` until we find one to satisfy the stated condition.

```{r threshold}

# Define p

p <- data$pred_eq

# Define the function

f <- function(p, x) {
  
  val1 <- round(sum(p))
  
  val2 <- sum(p >= x)
  
  out <- val1 - val2
  
  return(out)
  
}

# Minimise WRT x

out <- 1 # output of f
x <- 0.75 # first guess at x
x0 <- x
it <- 0  # iteration number

while(out != 0){
  
  inc <- 10e-5 # increment size
  
  out <- f(p, x) # get output of f
  
  mult <- 10^(floor(log10(sqrt(out**2)))-2) # multiplier to increment, depends
                                            # on how far we are from 0
  
  ifelse(out < 0,
         x <- x + (inc * mult),
         x <- x - (inc * mult))
  
  it <- it + 1
  
  if(it %% 100 == 0){
    if(x == x0){ # if x unchanged in 100 iterations
      print(paste0("Best guess: x = ", x, " to accuracy ", inc))
      break
    }
   x0 <- x # record value of x
   print(x0)
  }
  
}

```




