---
title: "Model Diagnostic"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}

source(here::here("setup_environment.R")) # For libraries and filepaths

detach("package:tidylog", unload=TRUE)

c_scot <- read.xlsx(paste0(data_folder,
                           pub_day,
                           "/diagnostics/",
                           "c_stats.xlsx"),
                    sheet = "Scotland")

c_pdg <- read.xlsx(paste0(data_folder,
                          pub_day,
                          "/diagnostics/",
                          "c_stats.xlsx"),
                   sheet = "Diagnosis")

c_loc <- read.xlsx(paste0(data_folder,
                          pub_day,
                          "/diagnostics/",
                          "c_stats.xlsx"),
                   sheet = "Location")
```

## Model Performance

### ROC, AUC and C-statistic

The [ROC](https://www.statology.org/interpret-roc-curve/) (Receiver Operating Characteristic) curve is used to show how well the model discriminates between outcomes. It plots the true positive rate (sensitivity) against the false positive rate (1 - specificity) of the model. **For the HSMR model, the positive (1) outcome is death within 30 days of admission.**

A perfect model will have an ROC curve that reaches the top left corner of the axes [0, 1] to indicate a sensitivity and specificity of 1. The closer the curve is to this point, the better the discriminatory power of the model.

A poor curve would be close to the $y=x$ line - this is equivalent to a random guess, ie. the model is correct half of the time.

[AUC](https://www.statology.org/what-is-a-good-auc-score/) (Area Under the Curve) is a single value metric to describe the ROC curve.

Because we know that a perfect curve will have an AUC of 1, we know that **an AUC close to 1 suggests better discrimination**.

An AUC of 0.5, the area beneath the $y=x$ line, is equivalent to a random guess.

For a regression model that aims to predict risk, such as the HSMR model, we can also use the ROC curve to determine the [C-statistic](https://www.statology.org/how-to-interpret-the-c-statistic-of-a-logistic-regression-model/). **The C-statistic is equal to the AUC.**

If we take all possible pairs of patients, where Patient A died within 30 days of admission and Patient B did not, the C-statistic tells us what fraction of those pairs would see Patient A given a higher predicted risk than Patient B. It is a measure of how well the model is able to rank patients in order of risk.

AUC and ROC are calculated and plotted using [pROC](https://rdrr.io/cran/pROC/man/pROC-package.html).

There is no strict rule about what constitutes a good or bad C-statistic; however, for the HSMR model we use the following system.

| C-Statistic Value    | Discrimination Score |
|----------------------|----------------------|
| c $\leq$ 0.5         | Very Poor            |
| 0.5 $<$ c $\leq$ 0.6 | Poor                 |
| 0.6 $<$ c $\leq$ 0.7 | Fair                 |
| 0.7 $<$ c $\leq$ 0.8 | Good                 |
| 0.8 $<$ c $\leq$ 0.9 | Very Good            |
| c $>$ 0.9            | Excellent            |


### ROC for Scotland

The ROC curve for the logistic regression model is shown below.

```{r roc_curve, echo=FALSE, out.width = '80%'}

png_path <- paste0(data_folder,
                   pub_day,
                   "/diagnostics/",
                   "roc.png")

knitr::include_graphics(png_path)
```

The AUC, or C-statistic, for the whole of Scotland is `r round(c_scot$c_stat, 3)`. This means that the model is able to correctly discriminate between high-risk and low-risk patients `r round(c_scot$c_stat, 3)*100`% of the time.

We expect the C-statistic for the whole of Scotland to be greater than 0.9 (Excellent).

### C-Statistic By Hospital

We can also calculate the C-statistic for each named hospital in the publication. This lets us know if there are any locations which are particularly impacted by poor discrimination from the model.

**If the HSMR value for a particular hospital is outside of the warning limits, the first thing to check is how well the model predicts outcomes there.**

The full list of hospital-level C-statistics can be found in the `diagnostics` folder. A summary is shown below.

```{r c_hosp, echo=FALSE}

c_loc_sum <- c_loc %>%
  group_by(discrimination) %>%
  summarise('N Hospitals' = n(),
            'N Patients' = sum(pats)) %>%
  rename('Discrimination Score' = discrimination)

c_loc_sum$`Discrimination Score` <- factor(c_loc_sum$`Discrimination Score`,
                                           levels = c("Very Poor",
                                                  "Poor",
                                                  "Fair",
                                                  "Good",
                                                  "Very Good",
                                                  "Excellent",
                                                  NA))

knitr::kable(c_loc_sum %>% arrange(`Discrimination Score`),
             caption = "Summary of C-statistics by hospital.")
```
An `NA` C-statistic means that there was only 1 outcome for a particular hospital (eg. no deaths within 30 days), so the C-statistic could not be calculated.

Hospitals, if any, that showed **Fair or worse** discrimination are listed below.

```{r poor_loc, echo=FALSE, message=FALSE}
c_loc_poor <- c_loc %>% 
             filter(c_stat <= 0.7) %>% 
             arrange(c_stat) %>%
             select(Hospital = location_name,
                    `N Patients` = pats,
                    `C Statistic` = c_stat,
                    `Score` = discrimination)

if(nrow(c_loc_poor) > 0){
  knitr::kable(c_loc_poor,
             caption = "Hospitals with C-statistics below 0.7.")
}

```

### C-Statistic By Primary Diagnosis

Similarly, we can calculate the C-statistic for each primary diagnosis group.

The full list of diagnosis-level C-statistics can be found in the `diagnostics` folder. A summary is shown below.

```{r c_diag, echo=FALSE}

c_pdg_sum <- c_pdg %>%
  group_by(discrimination) %>%
  summarise('N Diagnoses' = n(),
            'N Patients' = sum(pats)) %>%
  rename('Discrimination Score' = discrimination)

c_pdg_sum$`Discrimination Score` <- factor(c_pdg_sum$`Discrimination Score`,
                                           levels = c("Very Poor",
                                                  "Poor",
                                                  "Fair",
                                                  "Good",
                                                  "Very Good",
                                                  "Excellent",
                                                  NA))

good_plus <- c_pdg_sum %>%
  filter(`Discrimination Score` %in% c("Good", "Very Good", "Excellent")) %>%
  summarise(n_pdg = sum(`N Diagnoses`),
            n_pats = sum(`N Patients`)) %>%
  mutate(tot_pdg = nrow(c_pdg),
         tot_pats = sum(c_pdg$pats),
         pcent_pdg = 100*(n_pdg/tot_pdg),
         pcent_pats = 100*(n_pats/tot_pats))

knitr::kable(c_pdg_sum %>% arrange(`Discrimination Score`),
             caption = "Summary of C-statistics by primary diagnosis group.")
```

An `NA` C-statistic means that there was only 1 outcome for a particular diagnosis (eg. no deaths within 30 days), so the C-statistic could not be calculated.

`r round(good_plus$pcent_pdg, 2)`% of primary diagnosis groups, constituting `r round(good_plus$pcent_pats, 2)`% of patients, had Good, Very Good or Excellent C-statistics.

Diagnoses, if any, that showed **Fair or worse** discrimination are listed below.

```{r poor_pdg, echo=FALSE, message=FALSE}
knitr::kable(c_pdg %>% 
             filter(c_stat <= 0.7) %>% 
             arrange(c_stat) %>%
             select(`Diagnosis Code` = pdiag_grp,
                    `Diagnosis` = primary_diagnosis,
                    `N Patients` = pats,
                    `C Statistic` = c_stat,
                    `Score` = discrimination),
             caption = "Primary diagnosis groups with C-statistics below 0.7.")
```


