---
title: "Model Checks"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}

# WIP: revised version of the model checks markdown in line with old SSPS code.

library(pROC) # For ROC/AUC

source(here::here("setup_environment.R")) # For libraries and filepaths

detach("package:tidylog", unload=TRUE)

data <- read.csv(paste0(data_folder, # base files with predprob
                        pub_day,
                        "/base_files/", 
                        pub_day, 
                        "_SMR-with-predprob.csv"))

pdg <- read.csv(paste0(here::here(), # diagnosis lookup table
                       "/reference_files/",
                       "diag_grps_lookup_updated.csv")) %>%
  select(pdiag_grp = DIAGNOSIS_GROUP,
         primary_diagnosis = label)

data <- data %>%
  left_join(pdg)

```

context of some kind

## ROC, AUC and C-Statistic

The [ROC](https://www.statology.org/interpret-roc-curve/) (Receiver Operating Characteristic) curve is used to show how well the model discriminates between outcomes.

> For the HSMR model, the positive (1) outcome is death within 30 days of admission.

The ROC curve plots the true positive rate (sensitivity) against the false positive rate (1 - specificity) of the classifier. 



https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6465431/







A perfect model will have an ROC curve that reaches the top left corner of the axes [0, 1]. The closer the curve is to this point, the better the discriminatory power of the model. A poor curve would be close to the *y=x* line - this is equivalent to a random guess, ie. the model is correct half of the time.

AUC (**A**rea **U**nder the **C**urve) is a single value metric to describe the ROC curve. Because we know that a perfect curve will have an AUC of 1, we know that **an AUC close to 1 suggests better discrimination**. An AUC of 0.5, the area beneath the *y=x* line, is equivalent to a random guess.

For a regression model that aims to predict risk, such as the HSMR model, we can also use the ROC curve to determine the **C-statistic**. The C-statistic is equal to the AUC.

If we take all possible pairs of patients, where Patient A died within 30 days of admission and Patient B did not, the C-statistic tells us what fraction of those pairs would see Patient A given a higher predicted risk than Patient B. It is a measure of how well the model is able to rank patients in order of risk.

AUC and ROC are calculated and plotted using [pROC](https://rdrr.io/cran/pROC/man/pROC-package.html).

```{r cstats, echo=FALSE, message=FALSE}

c_stats <- data %>%         # get a dataframe of diagnosis codes only
  group_by(pdiag_grp, primary_diagnosis) %>%
  summarise(pats = n()) %>% 
  mutate(c_stat = NA,
         pcent_pats = 100*pats/sum(pats))       # add a dummy column for C-statistic

for(i in 1:nrow(c_stats)){                          # for every diagnosis code
  x <- data %>% filter(pdiag_grp == i)              # filter for this code
  if(length(unique(x$death30)) == 2){               # make sure both outcomes are represented
    auc_x <- as.numeric(auc(x$death30, x$pred_eq))  # get the AUC/C-statistic
    c_stats[i, 4] <- auc_x                          # write to dataframe
  }
  print(i)
}

```

```{r roc, echo=FALSE, message=FALSE}

roc_hsmr <- roc(data$death30, data$pred_eq)

auc_hsmr <- auc(roc_hsmr)

ggroc(roc_hsmr, 
      legacy.axes = TRUE,
      colour = '#3f3685') +
  theme_bw() +
  geom_abline(slope = 1,
              intercept = 0,
              colour = 'grey',
              linetype = 'dashed') +
  labs(title = paste0("ROC Curve - ", pub_day, " Publication")) +
  xlab("False Positive Rate") +
  ylab("True Positive Rate") +
  annotate("label", label=paste0("AUC = ", round(auc_hsmr, 3)), x = 0.9, y= 0.1)
```