---
title: "Check Table 1"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    code_folding: hide
    toc: yes
---

<style type="text/css">
.main-container {
  max-width: 72em;
  margin-left: auto;
  margin-right: auto;
}
</style>

# Setup

```{r message=FALSE}

# Load environment file
source(here::here("setup_environment.R"))

# Tidylog messages make notebook output messy
detach("package:tidylog", unload = TRUE)

library(glue)      # for glue()
library(DT)        # for tables

source("checking_functions.R")

```

```{r}

get_t1_data = function(date_str, sheet_name = "Raw Data"){
  
  path = glue("{data_folder}{date_str}/output/{date_str}-Table1-HSMR.xlsx")
  
  data = 
    read.xlsx(path, sheet = sheet_name) %>%
    select(deaths, pred, pats, smr, crd_rate, location_type, location, 
           location_name)
  
  return(data)
  
}

t1_old = get_t1_data(previous_pub)
t1_new = get_t1_data(pub_day)

```

# Check Scotland

* smr should = 1
* deaths should = pred

```{r}

scot_data = filter(t1_new, location_name == "Scotland")

print(scot_data)

```


# Compare to previous publication

We do expect there to be differences from the previous release, since they cover different time periods - but any drastic differences may be cause for further investigation.

Get a dataframe with the change from the last publication:

```{r}

# Columns containing the labels
join_cols = c("location_type", "location", "location_name")

t1_diff = compare_df(t1_new, t1_old, join_by = join_cols, 
                     df1_name = "new", df2_name = "old")

```


## Patients

* Numbers should usually be in the same ballpark

```{r}

t1_diff_px = 
  t1_diff %>%
  select(any_of(join_cols), starts_with("pats"))

writeLines(glue("Max % diff: {max(t1_diff_px[, 'pats.diff_pc'])}
                 Min % diff: {min(t1_diff_px[, 'pats.diff_pc'])}"))

```

```{r}
make_change_table(t1_diff_px, measure = "pats", quarter_filter = "all")
```


## Deaths

* Numbers should usually be in the same ballpark

```{r}

t1_diff_deaths = 
  t1_diff %>%
  select(any_of(join_cols), starts_with("deaths"))

writeLines(glue("Max % diff: {max(t1_diff_deaths[, 'deaths.diff_pc'])}
                 Min % diff: {min(t1_diff_deaths[, 'deaths.diff_pc'])}"))

```

```{r}
make_change_table(t1_diff_deaths, measure = "deaths", quarter_filter = "all")
```
